---
title: "Network Measurements"
author: "Benjamin Ortiz Ulloa"
date: "3/27/2018"
output: 
    ioslides_presentation:  
        widescreen: true
        smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(igraph)
library(igraphdata)
library(ggplot2)
library(ggraph)
data("UKfaculty")
```

## Degree

```{r}
V(UKfaculty)$dgr <- degree(UKfaculty)
V(UKfaculty)$dgr_color <- V(UKfaculty)$dgr %>%
  sapply(function(x){
    pct <- (length(.)-5)/length(.)
    pct <- quantile(., pct)
    x >= pct
  })

set.seed(1234)
ggraph(UKfaculty, layout = 'drl' ) +
  geom_edge_link(colour = 'lightgrey') +
  geom_node_point(aes(fill = dgr_color), size = 4, shape = 21) +
  theme_void() +
  theme(legend.position = 'none')
```

**Degree:** The number of edges connected to a node

## Betweeness

```{r}
V(UKfaculty)$btw <- betweenness(UKfaculty)

V(UKfaculty)$btw_color <- V(UKfaculty)$btw %>%
  sapply(function(x){
    pct <- (length(.)-5)/length(.)
    pct <- quantile(., pct)
    x >= pct
  })

set.seed(1234)
ggraph(UKfaculty, layout = 'drl' ) +
  geom_edge_link(colour = 'lightgrey') +
  geom_node_point(aes(fill = btw_color), size = 4, shape = 21) +
  theme_void() +
  theme(legend.position = 'none')
```

**Betweeness:** Defined by the number of shortest paths going through an edge

## Closeness

```{r, warning = F}
V(UKfaculty)$cls <- closeness(UKfaculty)
V(UKfaculty)$cls_color <- V(UKfaculty)$cls  %>%
  sapply(function(x){
    pct <- (length(.)-5)/length(.)
    pct <- quantile(., pct)
    x >= pct
  })


set.seed(1234)
ggraph(UKfaculty, layout = 'drl' ) +
  geom_edge_link(colour = 'lightgrey') +
  geom_node_point(aes(fill = cls_color), size = 4, shape = 21) +
  theme_void() +
  theme(legend.position = 'none')
```

**Closeness:** Defined by the number of steps required to access every other vertex from a given vertex

## Hub Score

```{r}

V(UKfaculty)$hub <- hub_score(UKfaculty)$vector
V(UKfaculty)$hub_color <- V(UKfaculty)$hub  %>%
  sapply(function(x){
    pct <- (length(.)-5)/length(.)
    pct <- quantile(., pct)
    x >= pct
  })

set.seed(1234)
ggraph(UKfaculty, layout = 'drl' ) +
  geom_edge_link(colour = 'lightgrey') +
  geom_node_point(aes(fill = hub_color), size = 4, shape = 21)  +
  theme_void() +
  theme(legend.position = 'none')
```

**Hub Score:** Defined by the number of Authorities the node points to

## Authority Score

```{r}
V(UKfaculty)$auth <- authority_score(UKfaculty)$vector
V(UKfaculty)$auth_color <- V(UKfaculty)$auth  %>%
  sapply(function(x){
    pct <- (length(.)-5)/length(.)
    pct <- quantile(., pct)
    x >= pct
  })

set.seed(1234)
ggraph(UKfaculty, layout = 'drl' ) +
  geom_edge_link(colour = 'lightgrey') +
  geom_node_point(aes(fill = auth_color), size = 4, shape = 21) +
  theme_void() +
  theme(legend.position = 'none')
```

**Authority Score:** Defined by the number of Hubs pointing to the node

## Page Rank

```{r}
V(UKfaculty)$pr <- page_rank(UKfaculty)$vector
V(UKfaculty)$pr_color <- V(UKfaculty)$pr %>%
  sapply(function(x){
    pct <- (length(.)-5)/length(.)
    pct <- quantile(., pct)
    x >= pct
  })

set.seed(1234)
ggraph(UKfaculty, layout = 'drl' ) +
  geom_edge_link(colour = 'lightgrey') +
  geom_node_point(aes(fill = pr_color), size = 4, shape = 21) +
  theme_void() +
  theme(legend.position = 'none')
```

Google interprets a link from page A to page B as a vote, by page A, for page B ... Votes cast by pages that are themselves ‘important’ weigh more heavily and help to make other pages ‘important.’ 