---
title: 'Do It Yourself: Part 1'
author: "Benjamin Ortiz Ulloa"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, pull the trumpworld data using `read_graph`

```{r}
library(igraph)
trumpworld <- read_graph(file = '../Data/trumpworld.graphml', format = 'graphml')
```

What kind of graph is this? Is it directed or undirected? What are the vertex attributes? What are the edge attributes?

```{r}
#Enter code here



#press "play" once you're done
```

There are two vertex attributes? Which of these attributes would hold a value of "DONALD J. TRUMP"? Which one of these attributes would have an attribute of "Person"? Try to query the graph for a vertex that represents a "Person" named "DONALD J. TRUMP"

```{r}
#Hint:
#V(trumpworld)[ATTRIBUTE1 == 'Person'][[ATTRIBUTE2 == 'DONALD J. TRUMP']]



#press "play" once you're done
```

Now, try to find any "Organization" that has "DONALD J. TRUMP" in it's name.

```{r}
#Hint:
#grepl("DONALD J. TRUMP", ATTRIBUTE)
#will return TRUE for any index of ATTRIBUTE that contains "DONALD J. TRUMP" 
#and will return FALSE for any index of ATTRIBUTE that does not contain "DONALD J. TRUMP"



#press "play" once you're done
```

Create a Vertex attribute called name that equals the Vertex attribute id

```{r}
#Hint:
#the "$" and "<-" work well together



#press "play" once you're done.
```

"MICHAEL FLYNN" is the "Founder and CEO" of something. What is it?

```{r}
#Hint:
#E(trumpworld)["MICHAEL FLYNN" ??? V(trumpworld)][[ATTRIBUTE == "Founder and CEO"]]



#press 'play' once you're done.
```
