---
title: 'Do It Yourself: Part 3'
author: "Benjamin Ortiz Ulloa"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(igraph)
trumpworld <- read_graph(file = '../Data/trumpworld.graphml', format = 'graphml')
V(trumpworld)$name <- V(trumpworld)$id
```

Let's create a subgraph of vertices that does not include "DONALD J. TRUMP"

```{r}
#fill out the code below 

notTrump <-

#run the code when you're read  
```

Create a matrix that calculates the **distances** between all 'Person' type nodes. This should be derived by the `notTrump` graph calculated above. Replace any value that is greater than 5 with 0.

```{r}
#hint:
#MATRIX[MATRIX > 5] <- 0

from_names <- V(notTrump)[type == 'Person']
to_names <- V(notTrump)[type == 'Person']

peopleDistanceMatrix <- distances(notTrump, from_names, to_names)
peopleDistanceMatrix[peopleDistanceMatrix > 3] <- 0
```

Create an undirected graph from the `peopleDistanceMatrix` you created above and call the value of the edges "distance"

```{r}

notTrumpGraph <- graph_from_adjacency_matrix(peopleDistanceMatrix, 'undirected', weighted = 'distance')

```

Who has the most connections in the "notTrumpGraph"?

```{r}


```

Choose a centrality score and determine which person is deemed most central in the 'notTrumpGraph'

```{r}

```