---
title: "Part 3: Exploratory Functions"
output:
  html_document:
    df_print: paged
  html_notebook: default
---

Ok, so we can now access information inside nodes and edges, as well as create subgraphs. What else can we do with these objects?

# Exploratory functions

Other than querying the nodes and edges directly, igraph provides a number of functions to explore graphs. igraph is a very large package that offers a lot of functionality, but here we will focus on 

* Functions that provide information specific to each component
* Functions that provide information on how each component relates to the overall graph

### Setup

```{r message = F}
library(igraph)
library(igraphdata)

data("UKfaculty")

UKfaculty
```

If we examine the graph, we know that this is both a directed and weighted graph with the vertex attribute of **Group** and the edge attribute of **weight**. Let's take a look at these attributes.

```{r}
UKfaculty %>%
  V() %>%
  .$Group %>%
  summary()

UKfaculty %>%
  E() %>%
  .$weight %>%
  summary()
```

```{r}
UKfaculty %>%
  V() %>%
  .$Group %>%
  unique() %>%
  sort()

UKfaculty %>%
  E() %>%
  .$weight %>%
  unique() %>% 
  sort()
```

finally, let's make a quick visualization of this graph

```{r}
set.seed(4321)
plot(UKfaculty,
     vertex.label = V(UKfaculty)$Group,
     edge.curved = T,
     edge.arrow.size = .3)
```

We will be filtering this graph down into a subgraph to show the power of some of these functions. In order to keep things consistent between the graph and the subgraph, we will need to name the vertices.

```{r}
V(UKfaculty)$name <- paste('person', V(UKfaculty), sep = '_')
```
# Functions that provide information specific to the components

## Functions that provide information specific to edges

Sometimes we know the edges we want, but don't know the nodes they are connected to. This often happens when we query the edges for a specific attribute. We can get get this information. 

Let's look at the relationships that have strong weights. Let's look for the number of quality connections. The above analysis showed the weight's 3rd quartile is 6. So we will use this as the criteria to define a relationship as strong.

For directed edges, we can find the node that is at the head (what the edge is *pointing to*) by using the `head_of()` function. For undirected graphs, the head is simply the second node mentioned in the edge description.

```{r}
topLiked <- UKfaculty %>%
  E() %>%
  .[weight >= 6] %>%
  head_of(UKfaculty, .)

topLiked %>% sort

hist(topLiked, 
     vcount(UKfaculty),
     col = 'lightblue')

topLiked$name %>%
  table %>%
  sort(T) %>%
  head()
```

Alternatively, we can get the nodes at the tail of a directed edge (where the edge is *coming from*) by using the `tail_of()` function. For undirected graphs, the tail is simply the first node mentioned in the edge description.

```{r}
topLikers <- UKfaculty %>%
  E() %>%
  .[weight >= 6] %>%
  tail_of(UKfaculty, .)

hist(topLikers, 
     vcount(UKfaculty), 
     col = 'lightblue')

topLikers$name %>%
  table() %>%
  sort(T) %>%
  head()
```

You can also get a matrix of vertex ids. The first column represents the tail of the edges and the second column represents the head of the edges.

```{r}
UKfaculty %>%
  E() %>%
  .[weight >= 6] %>%
  ends(UKfaculty, .) %>%
  head()
```
 
Let's go ahead and identify who the most liked person is and who likes the most people. Remember, we only care about edges that have a weight of greater than or equal to 6.The results would be different if we included all edges.
 
```{r}
mostLikedPerson <- topLiked$name %>%
  table() %>%
  sort(T) %>%
  .[1] %>%
  names()

mostLikedPerson

likesTheMost <- topLikers$name %>%
  table() %>%
  sort(T) %>%
  .[1] %>%
  names() 

likesTheMost
```

## Functions that provide information specific to nodes

### ego()

The `ego()` function provides all the nodes connected to a node or nodes of interest. The connected nodes can either be directly connected or connected within a certain order of edges. Let's find all the people that liked the `mostLikedPerson` and let's also find all the people the person who `likesTheMost` likes.

```{r}
mlp_likedBy <- ego(UKfaculty, 1, mostLikedPerson, 'in')
mlp_likedBy

ltm_likes <- ego(UKfaculty, 1, likesTheMost, 'out')
ltm_likes
```

You can enter any number of vertices in the `ego()` function and the order of the list output will correspond to the order the vertices were entered. Here we will look for all connections (both incomming and outgoing) for each node.

```{r}
ego(UKfaculty, 1, c(mostLikedPerson, likesTheMost), 'all')
```

let's combine these two people together and see their subnetwork

```{r}
mlp_ltm <- c(mlp_likedBy[[1]], ltm_likes[[1]]) %>%
  induced_subgraph(UKfaculty, .)
mlp_ltm

set.seed(1234)
plot(mlp_ltm,
     vertex.color = sapply(V(mlp_ltm)$name, function(x){
       if(x %in% c(mostLikedPerson, likesTheMost)){
         'lightblue'
       } else {
         'lightgrey'
       }
     }),
     edge.curved = T,
     edge.arrow.size = .4,
     vertex.label = V(mlp_ltm)$Group)

```

`induced.subgraph`s include all edges between the included nodes, but we're just interested in the edges of interest. Particularly, we only want to see incoming edges for the `mostLikedPerson` and we only want to see outgoing edges for the person who `likesTheMost`. Let's add a filtering step. 

```{r}
mlp_ltm <- mlp_ltm %>%
  E() %>%
  .[likesTheMost %->% V(mlp_ltm) | mostLikedPerson %<-% V(mlp_ltm)] %>%
  .[weight >= 6] %>%
  subgraph.edges(mlp_ltm, .) 

mlp_ltm

set.seed(1234)
plot(mlp_ltm,
     vertex.color = sapply(V(mlp_ltm)$name, function(x){
       if(x %in% c(mostLikedPerson, likesTheMost)){
         'lightblue'
       } else {
         'lightgrey'
       }
     }),
     edge.curved = T,
     edge.arrow.size = .4,
     vertex.label = V(mlp_ltm)$Group)
```

### degree()

A degree of a node is the simply the raw number of edges incident to it. You can generate this number by using the `degree()` function and can distinguish what direction of edge you want to count.

```{r echo = T}
degree(UKfaculty, mode = 'out') %>%
  sort(T) %>%
  head()

degree(UKfaculty, mode = 'in') %>%
  sort(T) %>%
  head()

degree(UKfaculty, c(likesTheMost, mostLikedPerson)) %>%
  sort(T)
```

### strength() - Weighted Degree

Rather than getting the raw count of incident edges, we can also add the edge attributes stored in them. This is done by using the `strength()` function. This is valuable, because as we seen earlier, the raw number of edges doesn't necessarily show the strength of those relationships. Remember that the node with the most incoming edges doesn't match the person we identified as the person with the most quality connections. 

```{r echo=T}
strength(UKfaculty, mode = 'out', weights = E(UKfaculty)$weight) %>%
  sort(T) %>%
  head()
strength(UKfaculty, mode = 'in', weights = E(UKfaculty)$weight) %>%
  sort(T) %>%
  head()
strength(UKfaculty, c('person_31', 'person_29', 'person_69'),  weights = E(UKfaculty)$weight)
```

### Using Strength and Degree together

We can get the average weight of edges incident to a vertex by combining the information `strength()` and `degree()` functions together.

```{r}
strength(UKfaculty, mode = 'out') %>%
  { ./degree(UKfaculty, mode = 'out')} %>%
  sort(T) %>% 
  head

strength(UKfaculty, mode = 'in') %>%
  { ./degree(UKfaculty, mode = 'in')} %>%
  sort(T) %>% 
  head
```

# Functions that provide information on how each component relates to the overall graph

## Walks

### shortest_paths

The shortest path between two vertices is the path that uses the least edges. This can be caluculated by using the `shortest_paths()` function

```{r}
sp <- shortest_paths(UKfaculty, 
               likesTheMost, 
               mostLikedPerson, 
               'out', 
               output = 'both')
sp

UKfaculty %>%
  E() %>%
  .[[. %in% sp$epath[[1]]]]
```

It's important to note, however, that shortest paths use the weight attribute of edges. It also interpret the weights as euclidian distances - that is, paths with less accumulated weight are prioritized over paths with more accumulated weight. Sometimes, we want to ignore the weight attribute because it doesn't necessarily make sense in the context of our data.

```{r}

shortest_paths(UKfaculty,
               likesTheMost,
               mostLikedPerson,
               'out',
               output = 'both',
               weights = NA) %>%
  .$epath %>%
  .[[1]] %>%
  {E(UKfaculty)[[E(UKfaculty) %in% .]]}


```

### distances

If you don't care so much about the actual path and only want to know know the magnitude of the walks, then you can use the `distance()` function. Again, weight matters.

```{r}
distances(UKfaculty, c(likesTheMost , mostLikedPerson), c(likesTheMost, mostLikedPerson), 'out')
distances(UKfaculty, c(likesTheMost , mostLikedPerson), c(likesTheMost, mostLikedPerson), 'in')
distances(UKfaculty, c(likesTheMost , mostLikedPerson), c(likesTheMost, mostLikedPerson), 'all', weights = NA)
```

### Random Walks

From the given start vertex, take the given number of steps, choosing an edge from the actual vertex uniformly randomly. If the graph is directed, then the random walk only uses edges that traverse outwards. Random walks are fundamental in both community detection and centrality measurements. 

```{r}
set.seed(1234)
random_walk(UKfaculty, likesTheMost, 2)
```
```{r}
sapply(1:100, function(i){
  random_walk(UKfaculty, likesTheMost, 2)[[2]]
}) %>% 
  table() %>%
  sort(T)
```

## Centrality Scores

### betweenness()

The betweenness value is roughly defined as the number of shortest paths going through a vertex or an edge. A betweenness value of 0 means that no shortest paths go through that vertex - that vertex is not essential for connecting any two vertices. Weights are interpreted as distances and a path with a lower weight will be chosen over a path with a higher weight. The betweenness value of a vertex can be calculated with the `betweenness()` function.

```{r}
betweenness(UKfaculty, likesTheMost)
```

```{r echo = T}
V(UKfaculty)$betweenness <- betweenness(UKfaculty)
V(UKfaculty)[[betweenness == max(betweenness)]]
```

```{r}
V(UKfaculty)$unweightedBetweenness <- betweenness(UKfaculty, weights = NA)
V(UKfaculty)[[unweightedBetweenness == max(unweightedBetweenness)]]
```

Betweenness can also be calculated for edges by using the `edge_betweeness()` function:

```{r}
E(UKfaculty)$betweenness <- edge_betweenness(UKfaculty)
E(UKfaculty)[[betweenness == max(betweenness)]]
```
```{r}
E(UKfaculty)$unweightedBetweenness <- edge_betweenness(UKfaculty, weights = NA)
E(UKfaculty)[[unweightedBetweenness == max(unweightedBetweenness)]]
```

### closeness()

Closeness measures how many steps is required to access every other vertex from a given vertex.

```{r, warning=F}
closeness(UKfaculty, mode = 'out') %>% 
  sort(T) %>% 
  head

closeness(UKfaculty, mode = 'in') %>%
  sort(T) %>%
  head
```

### eigen_centrality()

In general, vertices with high eigenvector centralities are those which are connected to many other vertices which are, in turn, connected to many others (and so on). 

```{r}
eigen_centrality(UKfaculty, directed = F)$vector %>%
  sort(T) %>%
  head()
```

## authority_score() & hub_score() 

Keinberg's Hyperlink-Induced Topic Search (HITS) - also known as Hubs and Authorities - stemmed from a particular insight into the creation of web pages when the Internet was originally forming; that is, certain web pages, known as hubs, served as large directories that were not actually authoritative in the information that they held, but were used as compilations of a broad catalog of information that led users direct to other authoritative pages. In other words, a good hub represented a page that pointed to many other pages, and a good authority represented a page that was linked by many different hubs. These can be calculated using the `authority_score()` and `hub_score()` functions.

```{r}
authority_score(UKfaculty)$vector %>% 
  sort(T) %>%
  head()

page_rank(UKfaculty)$vector %>%
  sort(T) %>%
  head()
```

### page_rank()

PageRank is a centrality metric created by Google.PageRank works by counting the number and quality of links to a page to determine a rough estimate of how important the website is. The underlying assumption is that more important websites are likely to receive more links from other websites. For undirected graphs, PageRank treats each edge as a bidirectional edge. PageRank can be calculated by using `page_rank()`

```{r}
page_rank(UKfaculty, directed = F)$vector %>%
  sort(T) %>%
  head()

page_rank(UKfaculty, weights = NA, directed = F)$vector %>%
  sort(T) %>%
  head()

```



