---
title: 'Part 2: A Review of magrittr'
output:
  html_document: default
  html_notebook: default
---


# Introduction

In this section we will cover some essential **R** functions and packages that facilitate the creation and exploration of networks. We will look at both useful functions in *base R* as well as their **tidverse** equivalents:

1. Iteration:
    + apply
    + *apply and map_\*
    + mapply and map2_\*
2. string:
    + stringr
        + str_detect
        + str_replace
    + grep/grepl/gsub
3. tidyverse
    + un/nest()
    + _join
    + filter
4. magrittr
5. miscellaneous
    + t()
    
    
## Starwars Network

```{r message = F}
library(tidyverse)
library(magrittr)
library(igraph)
data(starwars)
```

```{r}
starwars
starwars$films[[1]]
starwars %>% select(name, films)
```

```{r}
sw_character_films <- starwars %>%
  select(name, films) %>%
  unnest() %>%
  rename(from = name,
         to = films) %>%
  mutate(type = 'APPEARED IN')

sw_character_films

sw_characters <- sw_character_films %>%
  select(from) %>%
  rename(name = from) %>%
  distinct() %>%
  mutate(type = 'CHARACTER')

sw_characters

sw_films <- sw_character_films %>%
  select(to) %>%
  rename(name = to) %>%
  distinct() %>%
  mutate(type = 'FILM') 

sw_films
```

```{r}
g <- graph_from_data_frame(sw_character_films, F, rbind(sw_characters, sw_films))
```

```{r}
plot(g,
     vertex.size = 5,
     vertex.color = sapply(V(g)$type,
                           function(x){switch(x, 'CHARACTER' = 'lightblue', 'FILM' = 'orangered')}),
     vertex.label.cex = .65
     )
```

## generalizing the network creation process 

```{r}
head(starwars)
```

```{r}
sw_net_creator <- tibble(
  from = 'name',
  to = c('hair_color', 'skin_color', 'eye_color', 'gender', 'homeworld', 
         'species', 'films', 'vehicles', 'starships'),
  type = c('HAS HAIR COLOR', 'HAS SKIN COLOR', 'HAS EYE COLOR', 'HAS GENDER', 'HAS HOMEWORLD',
           'HAS SPECIES', 'APPEARS IN', 'PILOTS (VEHICLE)', 'PILOTS (STARSHIP)')
)
```

```{r}
sw_el_creator <- function(from, to, type){
  df <- starwars %>%
    select(from, to) %>%
    unnest() %>%
    mutate(type = type) %>%
    rename_("from" = from,
           "to" = to)
}
```

```{r}
sw_el <- pmap(sw_net_creator, sw_el_creator) %>%
  bind_rows %>%
  filter(!is.na(to),
         to != "none")
```

```{r}
sw_nl <- c(sw_net_creator$from,
           sw_net_creator$to) %>%
  unique %>% 
  lapply(function(x){
    if(x == 'name'){
      tibble(
        name = starwars[[x]],
        type = 'CHARACTER',
        height = starwars[['height']],
        mass = starwars[['mass']],
        birth_year = starwars[['birth_year']]
        ) 
      } else {
        tibble(
          name = starwars[[x]] %>% unlist,
          type = ifelse(str_detect(x, 'color'), 'COLOR', str_to_upper(x))
        ) %>%
          filter(!is.na(name))
      }
    }) %>%
  bind_rows() %>%
  distinct() %>%
  filter(name != 'none')
```

```{r}
sw_g <- graph_from_data_frame(sw_el, T, sw_nl)
```

```{r}
sw_g %>%
  V(.) %>%
  .[str_detect(name, 'Naboo')] %>%
  {E(sw_g)[. %--% V(sw_g)]} %>%
  .['Padmé Amidala' %--% V(sw_g)] %>%
  .[type == 'PILOTS (STARSHIP)']
```

## making indirect connections

```{r}
sw_nodes <- as_data_frame(sw_g, 'vertices')
sw_edges <- as_data_frame(sw_g, 'edges')
head(sw_nodes)
head(sw_edges)
```

```{r}
sw_nested_films <- sw_edges %>%
  filter(type == 'APPEARS IN') %>%
  select(from, to) %>%
  rename(character = from) %>%
  nest(character)

sw_nested_films
```

```{r}
sw_chr_chr <- sw_edges %>%
  filter(type == 'APPEARS IN') %>%
  select(from, to) %>%
  left_join(sw_nested_films) %>%
  unnest()

sw_chr_chr %>%
  head
```

```{r}
sw_chr_chr %<>% 
  group_by(from, character) %>%
  summarise(shared_movies = n())

sw_chr_chr %>% 
  head
```

```{r}
sw_chr_chr %>% 
  arrange(desc(shared_movies)) %>%
  filter(from != character) %>%
  head
```

```{r}
sw_chr_chr %>% 
  arrange(desc(shared_movies)) %>%
  filter(from != character) %>%
  head() %>%
  apply(1, sort)
```

```{r}
sw_chr_chr %>% 
  arrange(desc(shared_movies)) %>%
  filter(from != character) %>%
  select(from, character) %>%
  head() %>%
  apply(1, sort) %>%
  t()
```

```{r}
sw_unique_chr_chr <- sw_chr_chr %>%
  select(from, character) %>%
  filter(from != character) %>%
  apply(1, sort) %>%
  t() %>%
  as.tibble() %>%
  distinct() %>%
  rename(from = V1, character = V2)
```

```{r}
sw_unique_chr_chr %<>%
  left_join(sw_chr_chr)

sw_unique_chr_chr %>%
  arrange(desc(shared_movies))
```

```{r}
sw_unique_chr_chr %<>%
  rename(to = character) %>%
  mutate(type = 'APPEARED WITH')
```

```{r}
sw_edges <- list(sw_edges, sw_unique_chr_chr) %>%
  bind_rows
```

```{r}
sw_g <- graph_from_data_frame(sw_edges, F, sw_nodes)
```

```{r}
plot(sw_g,
     vertex.size = 5,
     vertex.label.cex = .65)
```

```{r}
just_char <- sw_g  %>% 
  {. - E(.)[type != 'APPEARED WITH'] } %>%
  {. - V(.)[type != 'CHARACTER']}
just_char
```

```{r}
plot(just_char,
     vertex.size = 5, 
     vertex.label.cex = .6)
```

```{r}
just_char %>%
  {. - E(.)[shared_movies < 4]} %>%
  {. - V(.)[degree(.) == 0]} %>% 
  plot(
    .,
    vertex.size = 5,
    vertex.label.cex = .7
  )
```

```{r}
V(sw_g)$height <- as.double(V(sw_g)$height)
write_graph(sw_g, '../Data/starwars_knowledge_graph.graphml', 'graphml')
```

## Bonus for after next section
```{r}
vader_g <- sw_g - E(sw_g)[type == 'APPEARED WITH']

anakin_movies <- vader_g %>%
  E() %>%
  .[type == 'APPEARS IN'] %>%
  (function(x){
    vader <- x[V(vader_g)[str_detect(name, 'Vader')] %--% V(vader_g)] %>%
      head_of(vader_g, .)
    anakin <- x[V(vader_g)[str_detect(name, 'Anakin')] %--% V(vader_g)] %>%
      head_of(vader_g, .)
    anakin[!anakin %in% vader]
  })

vader_g <- vader_g %>% 
  add_edges(
   matrix(c(rep('Darth Vader', 2), anakin_movies$name),2, 2, T),
   type = 'APPEARS IN' 
   )
```

```{r}
get_indirect <- function(graph, interest, connector, weightType, edgeType){
  el <- as_data_frame(graph, 'edges')
  nl <- as_data_frame(graph, 'vertices')
  
  main <- el %>%
    filter(type == connector) %>%
    select(from, to)
  
  other <- main %>%
    rename_("changed1" = interest) %>%
    nest(changed1)
  
  main %<>%
    rename_('changed2' = interest) %>%
    left_join(other) %>%
    unnest() %>%
    group_by(changed1, changed2) %>%
    summarise(weight = n()) 
  
  unique_main <- main %>%
    select(changed1, changed2) %>%
    apply(1, sort) %>%
    t() %>%
    as.tibble() %>%
    distinct() %>%
    rename(changed1 = V1,
           changed2 = V2) %>%
    filter(changed1 != changed2) %>%
    left_join(main) %>%
    mutate(type = edgeType)
  
  names(unique_main) <- c('from', 'to', weightType, 'type')
  el <- list(el, unique_main) %>%
    bind_rows
  graph_from_data_frame(el, F, nl)
}

vader_g <- get_indirect(vader_g, interest= 'from', connector = 'APPEARS IN', weightType = 'shared_movies', edgeType = 'APPEARED WITH')
```

```{r}
vader_g %>%
 {. - E(.)[type != 'APPEARED WITH']} %>%
 {. - V(.)[type != 'CHARACTER']} %>%
 {. - E(.)[shared_movies < 4 ]} %>%
 {. - V(.)[degree(.) == 0]} %>%
  plot(
    ., 
    vertex.size = 5,
    vertex.label.cex = .7
  )
```


```{r message = F}
library(igraph)
library(magrittr)
```





## Quick Review of Magrittr:



```{r}
g <- make_empty_graph(directed = T) +
  vertices(c('a', 'b', 'c', 'd', 'e'),
           type = "letter",
           order = c(1, 2, 3, 4, 5)) +
  edges(c('a', 'b',
          'b', 'c',
          'b', 'd',
          'c', 'b',
          'c', 'e'),
        type = "connection",
        weight = c(1, 1, 2, 1, 2))
```

```{r}
V(g)
```

can also be written as

```{r}
g %>% V(.)
```

The `.` dictates where the left hand side goes. This `.` behaves just like the object on the left hand side would.

```{r}
g %>% 
  V(.) %>% 
  .$type
```

The right hand side utilize the left hand side in mulitple places

```{r}
g %>% 
  V(.) %>% 
  .$type %>% 
  paste(., 1:length(.), sep = '_')
```

The right hand side can even be an anonymous function

```{r}
g %>% 
  V(.) %>% 
  .$type %>%
  paste(., 1:length(.), sep = '_') %>% 
  (function(x){
    assign('savingStep', x, .GlobalEnv)
    
    x %>%
      gsub(1, 'first', .) %>%
      gsub(length(.), 'last', .)
  })
```

The beautiful thing about the anonymous function is that it allows us to save the current state of an object to reference later without preventing us from moving forward with our manipulation on it.  

```{r}
paste(V(g))
```